---
title: "Map"
author: "Halie"
date: "7/13/2021"
output: html_document
---

# 4. Map

## 4.1 Read in Spatial Data

### 4.1.1 Install packages

```{r}
# require() is like library() except returns FALSE if missing (vs error)
if(!require(librarian)){
  install.packages("librarian")
  library(librarian)
  library(Rcpp)
}

#librarian::shelf() is like library() except installs package if missing, even from Github if include owner/repo
shelf(NOAA-EDAB/ecodata, sf)

```

## 4.2 Get spatial data

```{r}
epu_sf <- ecodata::epu_sf %>%
  st_transform(4326)

```

So we see a geometry list column.

```{r}
class(epu_sf)

g1 <- epu_sf$geometry[1]
# see in Environment pane, expand g1

plot(epu_sf)
plot(epu_sf["EPU"])

shelf(mapview)

mapview(epu_sf)

shelf(leaflet)

leaflet() %>%
  #addTiles() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addPolygons(data = epu_sf)
```

## 3.3 Group by

sf is "tidy"

## 3.4 Extract from ERDDAP

* ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01 degrees, 2002-present ,Monthly- Data Access form (https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html)

* CoastWatch ERDDAP (https://coastwatch.pfeg.noaa.gov/erddap/index.html): search for "SST":
  + jplMURSST41mday: ERDDAP - Multi-scale Ultra-high Resolution (MUR) SST Analysis fv04.1, Global, 0.01°, 2002-present, Monthly - Data Access Form (https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html)

```{r}
shelf(here, rerddap)

sst_gd_rds <- here("data/sst_gd.rds")

epu_bb <- st_bbox(epu_sf)
epu_bb

sst_info <- info('jplMURSST41mday')
sst_info

if(!file.exists(sst_gd_rds)){
  sst_gd <- griddap(sst_info, fields = "sst", time = c("2020-06-16", "2021-06-16"),longitude = epu_bb[c("xmin", "xmax")], latitude = epu_bb[c("ymin", "ymax")])
  saveRDS(sst_gd,file  = sst_gd_rds)
}

sst_gd <- readRDS(sst_gd_rds)
sst_gd

names(sst_gd)
```

```{r}
shelf(dplyr, ggplot2, mapdata)

#coastline
coast <- map_data("worldHires", xlim = epu_bb[c("xmin", "xmax")], ylim = epu_bb[c("ymin", "ymax")], lforce = "e")

sst_df_last <- sst_gd$data %>%
  filter(time == max(time))

# summary(sst_last)

ggplot(data = sst_df_last, aes(x = lon, y = lat, fill = sst)) + geom_polygon(data = coast, aes(x = long, y = lat, group = group), fill = "grey80") + geom_tile() + scale_fill_gradientn(colors = rerddap::colors$temperature, na.value = NA) + theme_bw() + ylab("Latitude") + xlab("Longitude") + ggtitle("Latest SST")
```

```{r}
shelf(purrr, raster, sp, tidyr)
select <- dplyr::select

sst_tbl <- tibble(sst_gd$data) %>% 
  mutate(
    # round b/c of uneven intervals
    # unique(sst_gd$data$Lon) %>% sort() %>% diff() %>% table()
    lon = round(lon, 2), lat = round(lat, 2), date = as.Date(time, "%Y-%m-%dT00:00:00Z")) %>%
  select(-time) %>%
  filter(!is.na(sst)) # 13M to 8.8M rows

sst_tbl_mo <- sst_tbl %>%
  nest(data = c(lat, lon, sst)) %>%
  mutate(raster = purrr::map(data, function(x){
   #browser()
    sp::coordinates(x) <- ~ lon + lat
    sp::gridded(x) <- T
    raster::raster(x)
  }))

sst_stk <- raster::stack(sst_tbl_mo$raster)
names(sst_stk) <- strftime(sst_tbl_mo$date, "sst_%Y.%m")
raster::crs(sst_stk) <- 4326
```

```{r}
shelf(stringr)

epu_sst_avg <- raster::extract(sst_stk, epu_sf, fun = mean, na.rm = T)
epu_sst_sd <- raster::extract(sst_stk, epu_sf, fun = sd, na.rm = T)

epu_sst_tbl <- rbind(epu_sst_avg %>%
  as_tibble() %>%
  cbind(EPU = epu_sf$EPU, stat = "mean") %>%
  pivot_longer(-c(EPU, stat)),
epu_sst_sd %>%
  as_tibble() %>%
  cbind(EPU = epu_sf$EPU, stat = "sd") %>%
  pivot_longer(-c(EPU, stat))) %>%
  mutate(EPU = as.character(EPU), date = as.double(str_replace(name, "sst_", ""))) %>%
  select(-name) %>%
  pivot_wider(names_from = EPU, values_from = value)
```

```{r}
shelf(dygraphs)

epu_sst_tbl %>%
  filter(stat == "mean") %>%
  select(-stat) %>%
  dygraph()
```












